---
title: "Homework 2 Answers"
subtitle: "EPI 525"
date-modified: "today"
format: 
  html: 
    link-external-newwindow: true
    toc: true
  pdf: default 
editor_options: 
  chunk_output_type: console
---

These are the **some** numeric/short answers to the homework. Often, these answers are insufficient for your own work or solutions. I just wanted to give you a part of the answer to help guide you in the right direction.

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(gtsummary)
library(gt)
```

# Book exercises {.unnumbered}

## 2.6 Poverty and language {.unnumbered}  
  
The American Community Survey is an ongoing survey that provides data every year to give communities the current information they need to plan investments and services. The 2010 American Community Survey estimates that 14.6% of Americans live below the poverty line, 20.7% speak a language other than English (foreign language) at home, and 4.2% fall into both categories.

### a   {.unnumbered}

Not disjoint

### b     {.unnumbered}

```{r}
#| echo: false
#| output: false

.146-.042  # below poverty but don't speak foreign language
.207-.042 # speak foreign language at home but don't live below poverty

#first install venndiagram package.
#install.packages("VennDiagram") 
library(VennDiagram)
```

```{r}
#| echo: false
#| message: false
#| warning: false
#drawing diagram: 
grid.newpage()                                        
draw.pairwise.venn(
  area1 = .146,  # below poverty line
  area2 = .207, # speak foreign language at home
  cross.area = .042, # both
  category = c("below poverty line", "speak foreign language"),
  fill = c("red", "blue")
  )

```


### c     {.unnumbered}

```{r}
#| echo: false

diff <- (.146-0.042)*100
```


`r diff`% 

### d    {.unnumbered}

```{r}
#| echo: false

aorb <- ((.146+.207)-0.042)*100
```

`r aorb`%

### e     {.unnumbered}

```{r}
#| echo: false

neitherab <- 100 - aorb
```

`r neitherab`% 

### f     {.unnumbered}

Not independent

## 2.8 School absences {.unnumbered}

### a   {.unnumbered}

```{r}
#| echo: false

#define vars
d1 <- .25
d2 <- .15
d3 <- .28
d0 <- 1-d1-d2-d3
```

`r d0`

### b     {.unnumbered}

```{r}
#| echo: false

no_more_1 <- d0+d1
```

`r no_more_1`

### c     {.unnumbered}

```{r}
#| echo: false

at_least_one <- (1-d0)
```

`r at_least_one`

### d     {.unnumbered}

```{r}
#| echo: false

neither_sick <- (d0*d0)
```

`r neither_sick`

### e     {.unnumbered}

```{r}
#| echo: false

both_miss <- (1-d0)^2
```

`r both_miss`

## 2.10 Health coverage, frequencies {.unnumbered}  

```{r}
#| echo: false

#table. not required
(covg.tab <- data.frame(
  health_status = c("No","Yes","total"),
  Excellent = c(459,4198,4657),
  Very_good = c(727,6245,6972),
  Good = c(854,4821,5675),
  Fair = c(385,1634,2019),
  Poor = c(99,578,677),
  Total = c(2524,17476,20000)
)) %>% gt()
```

### a     {.unnumbered}

```{r}
#| echo: false

eh.no_covg <- (459/20000)
```

`r eh.no_covg`

### b     {.unnumbered}

```{r}
#| echo: false
a <- 4657/20000; b <- 2524/20000; c <- 459/20000
eh_or_no.hc <- a+b-c
```
`r eh_or_no.hc`


## 2.14 Health coverage, relative frequencies {.unnumbered}  

*The Behavioral Risk Factor Surveillance System (BRFSS) is an annual telephone survey designed to identify risk factors in the adult population and report emerging health trends. The following table displays the distribution of health status of respondents to this survey (excellent, very good, good, fair, poor) conditional on whether or not they have health insurance.*

```{r}
(covg.tab <- data.frame(
  health_status = c("No","Yes","total"),
  Excellent = c(0.0230, 0.2099,0.2329),
  Very_good = c(0.0364, 0.3123,0.3486),
  Good = c(0.0427,0.2410,0.2838),
  Fair = c(0.0192,0.0817,0.1009),
  Poor = c(0.0050,0.0289,0.0338),
  Total = c(0.1262,0.8738,1.0000)
)) %>% gt()
```
       
### a  

No

### b     {.unnumbered}

0.2329

### c    {.unnumbered}

```{r}
#| echo: false
eh <- 0.2329; hc <- 0.8738; eh_and_hc <- 0.2099
eh_give_hc <- eh_and_hc/hc
```

`r eh_give_hc`

### d    {.unnumbered}

```{r}
#| echo: false
no_hc <- 0.1262; eh_and_no_hc <- 0.0230
eh_give_no_hc <- eh_and_no_hc/no_hc
```

`r eh_give_no_hc`

### e     {.unnumbered}

No


## 2.18 Predisposition for thrombosis {.unnumbered}  

### a   {.unnumbered}

0.605

### b     {.unnumbered}

0.9997

## 2.24 Breast cancer and age {.unnumbered}  

### a  {.unnumbered}

```{r}
#| echo: false
#define variables
d30 <- 0.0044 #30-40
d40 <- 0.0147 #40-50
d50 <- 0.0238 #50-60
d60 <- 0.0356 #60-70
d70 <- 0.0382 #70-80
pos.g.d <- 0.85 #prob of positive test given disease
neg.g.dc <- 0.95 #prob of negative test given not having disease
pos.g.dc <- 1-neg.g.dc #prob of positive test given not having disease.
neg.g.d <- 1-pos.g.d
```


```{r}
#| echo: false

# Below we take advantage of R's capability of doing vector-wise calculations

# Create a vector of the prevalences
prevalence <- c(d30, d40, d50, d60, d70)

# Compute the PPV's using the vector of prevalences
PPV <- round((pos.g.d*prevalence) / (prevalence*pos.g.d + (1-prevalence)*pos.g.dc), digits = 3)
```
  
```{r}
#| echo: false

age_group <- c("30 - 40", "40 - 50", "50 - 60", "60 - 70", "70 - 80")

data.frame(age_group, prevalence, PPV) %>% 
  gt()
```

### b   {.unnumbered}

The PPV increased more with the higher specificity vs. the higher sensitivity.

## Non-book exercise

```{r}
#| echo: false
#name variables: 
a.g.d1 <- 0.80 # a given d1
a.g.d2 <- 0.90 # a given d2
a.g.d0 <- 0.01 # a given d0
d1 <- 0.009
d2 <- 0.001
d0 <- 0.99
```

```{r}
#| echo: false
#find the probability the patient has disease 2 given the have abdominal pain: 
PDA <- (a.g.d2*d2) / (a.g.d2*d2 + a.g.d1*d1 + a.g.d0*d0)
```

`r PDA`


