---
title: "Homework 9 Answers"
author: "Nicky Wakim"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    toc_float:
      collapsed: false
    number-sections: true
    number-depth: 3
    theme: lumen
    embed-resources: true
    html-math-method: mathjax
---

::: callout-warning
To see my math equations properly, you need to download the html file, then open it! One Drive does not show the math correctly!!
:::

```{r}
#| label: "setup"
#| include: false
knitr::opts_chunk$set(echo = TRUE)
```

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
```

# Book exercises {.unnumbered}


## 8.28 True or false, Part II {.unnumbered}

### (a)  {.unnumbered}

TRUE.

### (b)  {.unnumbered}

TRUE. 


### (c)  {.unnumbered}

FALSE.    

### (d)  {.unnumbered}

FALSE.    

## 8.34 Coffee and Depression {.unnumbered}

### (a)

### (b)-(f)
Instead of doing part (b) - (f), please run a hypothesis test using the Chi-squared test. 



### (g)

## 5.46 Child care hours {.unnumbered}



## 5.48 True/False: ANOVA, Part II {.unnumbered}

# R exercise

## Load all the packages you need below here. 

## R1: Palmer Penguins ANOVA


```{r}
library(palmerpenguins)
data(penguins)
```


### Dotplots

__Make a dotplot of the penguins' bill depths stratified by species type. Include points for the mean of each species type as well as a horizontal dashed line for the overall mean.__ _See example from class for the plot I'm describing._


### Technical conditions

__Investigate whether the technical conditions for using an ANOVA been satisfied. __

### Which groups significantly different?

__Based on the figure, which pairs of species look like they have significantly different mean bill depths?__ 


### Hypotheses in symbols or words

__Write out in symbols or words the null and alternative hypotheses.__  



### Run ANOVA in R

__Using R, run the hypothesis test and display the output.__  


### F statistic

__Using the values from the ANOVA table, verify (calculate) the value of the F statistic.__ 


### Decision?

__Based on the p-value, will we reject or fail to reject the null hypothesis? Why?__


### Conclusion

__Write a conclusion to the hypothesis test in the context of the problem.__

# Nonparametric-Tests


## NPT 1: (Wilcoxon) Signed-rank test

__Vegetarian diet and cholesterol levels__

When covering paired t-tests on Day 10 Part 2, the class notes used the example of testing whether a vegetarian diet changed cholesterol levels. The data are in the file `chol213.csv` at <https://niederhausen.github.io/BSTA_511_F23/data/chol213.csv>. In this exercise we will use non-parametric tests to test for a change and compare the results to the paired t-test.


### Hypotheses

__What are the hypotheses for the signed-rank test (2-sided) in the context of the problem?__ 

### Test in R

__Run the (Wilcoxon) Signed-rank test in R. What is the  p-value and how does it compare to the p-value of the sign test and the paired t-test (check the class notes for this)?__ 


## 8.38 (a) & (extra) Salt intake and CVD {.unnumbered}

Do **not** do parts (b)-(c) in the book

### (a) {.unnumbered}

* __You can use the expected cell counts from `expected()` in R (you do not need to compute them using the formula).__
* __Comment on whether the sample size condition is met or not for these data.__

### (extra) {.unnumbered}

__Run a Fisher's Exact test. Include the hypotheses and a conclusion in the context of the problem.__



# Extra R exercises (optional)

## R2: Palmer Penguins SLR


### Scatterplots

* __For each of the following pairs of variables, make a scatterplot showing the best fit line and describe the relationship between the variables.__
* In particular address 
  * whether the association is linear, 
  * how strong it is (based purely on the plot), and 
  * what direction (positive, negative, or neither).

1. __body mass vs. flipper length__



2. __bill depth vs. flipper length__



3. __bill depth vs. bill length__



### Correlations

* __For each of the following pairs of variables, find the correlation coefficient $r$.__

1. __body mass vs. flipper length__



2. __bill depth vs. flipper length__


3. __bill depth vs. bill length__


### Compare associations

 __Which pair of variables has the strongest association? Which has the weakest? Explain how you determined this.__


### Body mass vs. flipper length SLR

__Run the simple linear regression model for body mass vs. flipper length, and display the regression table output.__


### Regression equation

__Write out the regression equation for this model, using the variable names instead of the generic $x$ and $y$, and inserting the regression coefficient values.__

### Interpret intercept 

__Write a sentence interpreting the intercept for this example. Is it meaningful in this example?__ 


### Interpret slope

__Write a sentence interpreting the slope for this example.__ 








